### 2014-01-27 DGE_paper.tex
minor update

Marek's disease is an economically significant chicken disease that affects a poultry industry worldwide.
Total of \$2 billion loss has been reported from outbreaks.
The disease is caused by highly oncogenic Marek's disease virus (MDV) which causes T-cell lymphoma.
Vaccination is effective in reducing incidence of tumor formation; however, current MD vaccines do not
prevent infection or horizontal spread of the virus.
Improper use of vaccines is a key factor that drives evolution of highly virulent strains in vaccinated
flocks.
***Need to add intro about breed selection against infection***

Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
the disease. For example, chickens with MHC allele B$^{21}$ is highly resistant in contrast to chickens with
B$^{16}$ allele.
However, chickens line 6 and 7 , even with common allele (B$^6$), exhibit different phenotypic response to
MD.
The major unanswered question is what genetic factors contribute to susceptibility and resistance of
the disease.
***Need to add intro about gene and isoforms in immune system***
In this study, we identified several annotated and unannotated genes and isoforms that are differentially
expressed in response to MD infection. Pathway analysis and gene ontology suggest that these
genes may have important roles in immune response to MD.

@@

\subsection*{Custom Gene Annotations}
A large number of reads mapped to intergenic and intronic regions (table 1, supp.)
regions of Ensembl annotations (Galgal4.72)
indicates that many genes and isoforms are not included in the annotations.
Therefore, to extensively study gene and isoform expression, we employed \emph{de novo}
transcriptome assembly as well as genome-based assembly to build custom gene annotations (see methods).
The number of genes and isoforms is shown in Table \ref{tab:gene_models}.
Genes were identified by aligning against all vertebrate transcripts in GenBank database.
We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
as well as genes found in other organisms (Table \ref{}).

@@

Pathway analysis shows that genes in line 7 are significantly enriched (p $< 0.05$) in pathways involved in
immune system such as Toll-like receptor signaling, intestinal immune network for IgA production,
and Phagosome pathway, which are not significantly enriched in line 6 (fig. \ref{line67_kegg}).
Figure \ref{KEGG_phagosome} and \ref{KEGG_IgA} shows a diagram of phagosome and intestinal immune network
for IgA production repectively. Genes from line 6 and line 7 are highlighted with yellow and red color.
Common genes are highlighted with orange color.
These two pathways show that both line 6 and line 7 regulated MHC class I gene in response to infection,
but only line 7 regulated MHC class II gene.
It has been shown that Marek's disease virus (MDV), unlike other herpes viruses, up-regulates MHC class II gene
during infection. Moreover, regulation is believed to be mediated directly by MDV infection\cite{Niikura:2007fq}.
However, from these results, it appears that only MHC class II gene in line 7 was up-regulated.

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection{Gene Models Construction}
With the lack of high quality gene models for chickens. We employed two different approaches to
construct gene models from RNA-Seq reads.
First, short reads were quality trimmed with conditri/2.1\cite{} \texttt{(-cutfirst 10)}
and assembled using Velvet/1.21\cite{} and Oases/0.2.06\cite{} to obtain long transcripts.
Assembly was done with k value ranges from 21 to 31 for both local and global assembly
(described in Gimme paper\cite{}).
Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean\cite{}.
All transcripts were then aligned to chicken reference genome (galGal4, with short
contigs removed) with BLAT\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
Filtered alignments from BLAT were then used to produced gene models using Gimme.

Second, short reads were aligned to reference genome using Tophat/2.0.5\cite{} and mapping output were used to
build gene models by Cufflinks2\cite{} with default parameters.
Both set of gene models were combined using Gimme to build more comprehensive models
to be used in this study.

>>>>

Marek's disease is an economically significant chicken disease that affects a poultry industry worldwide.
Total of \$2 billion loss has been reported from outbreaks.
The disease is caused by highly oncogenic Marek's disease virus (MDV) which causes T-cell lymphoma.
Vaccination is effective in reducing incidence of tumor formation; however, current MD vaccines do not
prevent infection or horizontal spread of the virus.
Improper use of vaccines is a key factor that drives evolution of highly virulent strains in vaccinated
flocks.

***Need to add intro about breed selection against infection***

Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
the disease. For example, chickens with MHC allele B$^{21}$ is highly resistant in contrast to chickens with
B$^{16}$ allele.
However, chickens line 6 and 7 , even with common allele (B$^6$), exhibit different phenotypic response to
MD.
The major unanswered question is what genetic factors contribute to susceptibility and resistance of
the disease.

***Need to add intro about gene and isoforms in immune system***

In this study, we identified several annotated and unannotated genes and isoforms that are differentially
expressed in response to MD infection. Pathway analysis and gene ontology suggest that these
genes may have important roles in immune response to MD.

@@

\subsection*{Custom Gene Annotations}
A large number of reads mapped to intergenic and intronic regions (table 1, supp.)
regions of Ensembl annotations (Galgal4.72)
indicates that many genes and isoforms are not included in the annotations.
To extensively study gene and isoform expression, we employed \emph{de novo}
transcriptome assembly and a reference-guided assembly to build custom gene annotations (see methods).
The number of genes and isoforms is summarized in Table \ref{tab:gene_models}.
Genes were identified by aligning against non-redundant protein database from NCBI.
We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
as well as genes found in other organisms (Table \ref{}).

@@

\subsection*{Gene network analysis}
Pathway analysis shows that genes in line 7 are significantly enriched (p $< 0.05$) in pathways involved in
immune system such as Toll-like receptor signaling, intestinal immune network for IgA production,
and Phagosome pathway, which are not significantly enriched in line 6 (fig. \ref{line67_kegg}).
Figure \ref{KEGG_phagosome} and \ref{KEGG_IgA} shows a diagram of phagosome and intestinal immune network
for IgA production repectively. Genes from line 6 and line 7 are highlighted with yellow and red color.
Common genes are highlighted with orange color.
These two pathways show that both line 6 and line 7 regulated MHC class I gene in response to infection,
but only line 7 regulated MHC class II gene.
It has been shown that Marek's disease virus (MDV), unlike other herpes viruses, up-regulates MHC class II gene
during infection. Moreover, regulation is believed to be mediated directly by MDV infection\cite{Niikura:2007fq}.
However, from these results, it appears that only MHC class II gene in line 7 was up-regulated.

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection{Gene Models Construction}
Due to lack of complete gene models for chickens, we employed two methods to
construct gene models from RNA-Seq reads.
First, short reads were quality trimmed with conditri/2.1\cite{} \texttt{(-cutfirst 10)}
and assembled using Velvet/1.21\cite{} and Oases/0.2.06\cite{} to obtain long transcripts.
Assembly was done with k value ranges from 21 to 31 for both local and global assembly
(described in Gimme paper\cite{}).
Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean\cite{}.
All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
chromosomes removed) with BLAT\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
Filtered alignments from BLAT were then used to produced gene models using Gimme\cite{}.

Second, short reads were aligned to reference genome using Tophat/2.0.5\cite{} and 
gene models were built by Cufflinks2\cite{} with default parameters.
A combination of gene models from both set of gene models were used in this study.
